Expt 5: Mix of PD and BoS with spikes and two types of PD trials

### 2024-10-29 code/analysis_oct2024.Rmd
add analysis expt 5 prelim

---
title: "Mix of PD and BoS"
output:
  html_document: 
    toc: true
---

* 32 games each condition (chat, no-chat)
* humanness ratings are okay (75-77% -- in the acceptable range, probably lower than ideal)
* seeing very little effect of chat, probably because people don't use the chat very much (and group to group variation) 
* unlike prior pilots, not seeing an effect of chatted other times but not this one
  * could be chance / slightly different participant populations
  * could be different strategy space -- we did take away the points-counter which may make it harder to use that as an agreement point
  * could be this task is too easy
* about half of all utterances contain a color word 
* strategy wise, a lot of talk about fairness, also some about having an idea of a strategy but partner not getting it

so...what now:
* could give up on this project 
* one big issue for both design and interpretation is that people's incentives are hard to control or predict and aren't the numbers we put out (strong values of fairness, for instance)
  * the idea was to try to leverage game theory predictions, but people don't act like game
  theory agents / the payoffs that a person is choosing over aren't what we're displaying
  * possibly other areas of game theory would be better? 
* could force chat usage (someone must send message before selection can occur -- heavy handed, but could get us ceiling) 
* could up the difficulty -- more game types mixed together
* could switch back to offering running total of reward

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"), fig.width=7, fig.height=3)
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

@@

## Data import constants
data_location <- "data/new_study1_240520"

date_start <- lubridate::ymd("2024-05-20")

@@

@@

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
# #
#cost <- read_csv(here("player_payments.csv")) |> select(id, bonus) |> filter(bonus>0) |> write_csv(here("for_prolific.csv"))
# #
# sum(cost$cost)
```

```{r, include=F, eval=F}
d.treatments <- read_csv(here(data_location, "treatments.csv")) %>%
  rename(treatmentId = `_id`) %>%
  select(treatmentId, name) |>
  mutate(
    chat_cond = ifelse(str_detect(name, "no_chat"), "no_chat", "chat"),
    game_cond = "may2024"
  )

@@

d.complete <- d.round_results.raw %>%
  group_by(gameId, playerId) %>%
  summarize(max_rep = max(repNum)) %>%
  filter(max_rep == 39) %>%
  select(gameId, playerId) %>%
  write_csv(here(data_location, "complete.csv"))

@@

@@

exit %>%
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes))
```

Yay, they mostly think they're playing with a human!!

Looks like we have 32 games in each condition (from less attrition than I anticipated).

@@

ggplot(bonuses, aes(x = chat_cond, y = payoff, color = str_c(chat_cond))) +
  geom_jitter(width = .1, alpha = .1) +
  facet_grid(. ~ gametype) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )
```
Not seeing much of a difference in either category...

@@

game_chat |> filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(gameId, gametype) %>%
  summarize(pct_chat = mean(is.chat)) |> 
  ggplot(aes(x=reorder(gameId, pct_chat, mean), y=pct_chat, color=gametype))+geom_point()+
  theme(axis.text.x=element_blank())+
    coord_cartesian(ylim=c(0,1))

game_chat |> filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(gameId) %>%
  summarize(pct_chat = mean(is.chat)) |> 
  arrange(pct_chat) |> 
  mutate(x=1/n()*row_number()) |> ggplot(aes(x=x, y=pct_chat))+geom_point()+
    coord_cartesian(ylim=c(0,1))+geom_hline(yintercept=.1)+geom_hline(yintercept=.5)+labs(title="Game 4")

looks like people are more likely to talk in BoS than PD games. But there's not a lot of talking at all. Driven by group-differences -- some pairs never or almost never chatted even if they could -- others used chat a substantial fraction of the time.

@@

```{r}
ggplot(game_chat, aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(. ~ gametype) +
  geom_smooth(method = "lm")
```

@@

overall_talk <- game_chat |> left_join(group_talk)
ggplot(overall_talk |> filter(chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  geom_jitter(alpha = .1) +
  geom_smooth(method = "lm") +
  facet_grid(. ~ gametype) +
  theme(legend.position = "bottom")
```

Looks like talking on *other* trials might help in BoS if you didn't talk on this specific trial? Maaaybe? But there's a real data-sparsity issue.

@@

Near chance if you can't talk, above chance if you could but don't,  far above chance if you do.

@@

## PD

@@

Can get a reasonable option no matter what. 
```{r}
outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = repNum, y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot", geom="point", alpha=.5) +
  facet_grid(.~str_c(chat_cond, "_", is_chat))+
  geom_smooth(method="lm")+
  geom_hline(aes(yintercept = .5)) 
```

@@

```{r}
pair_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, playerId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    is_chat = ifelse(words > 0, 1, 0)
  ) |>
  group_by(gameId, gametype, chat_cond, repNum) |>
  summarize(how_many_chat = sum(is_chat)) |>
  group_by(gameId, gametype, chat_cond) |>
  summarize(pct_both = mean(how_many_chat == 2)) |>
  filter(chat_cond == "chat")

ggplot(pair_chat, aes(x = gametype, y = pct_both)) +
  geom_jitter(width = .2) +
  stat_summary(fun.data = "mean_cl_boot", color = "red")
```
Even in games that talk, there aren't *that* many trials where both people talk?

# Models

@@

```{r}
perf_priors <- c(
  set_prior("normal(5,2)", class = "Intercept"),
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 2)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

@@

```{r}
for_log_words <- for_mod |>
  filter(words > 0) |>
  mutate(log_words = log(words))

@@

```{r}
perf_word_mod <- brm(
  payoff ~ log_words * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_log_words,
  prior = perf_priors,
  file = here(model_location, "perf_words_expt1"),
  control = list(adapt_delta = .99)
)

@@

```{r}
exit <- read_csv(here(data_location, "exit.csv")) |> select(chat_cond, strategy) |> filter(!is.na(strategy))

@@

```{r}
color_words <- c("red","blue", "purple", "green","yellow", "brown", "pink", "violet","cyan", "grey", "peach", "purplish", "beige", "orange")
agreement_words <- c("okay", "ok", "yes", "sure")
color_regex <- str_c(color_words, collapse="|")
agreement_regex <- str_c(agreement_words, collapse="|")

@@

>>>>

---
title: "Expt 5: Mix of PD and BoS with spikes and two types of PD trials"
output:
  html_document: 
    toc: true
---

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"), fig.width=7, fig.height=3)
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(viridis)
library(knitr)
theme_set(theme_bw())

@@

## Data import constants
data_location <- "data/study5_241028"

date_start <- lubridate::ymd("2024-10-26")

@@

Note: one participant wasn't getting a partner, so they opened up a new tab, fiddled with the id and played themselves. 
So we exclude the game with player id YSB8RYRgF2tQjym2e, which is game AQXAv4FKrZxBrTQgE

@@

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) 
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
# #
# cost <- read_csv(here(data_location,"player_payments.csv"))
# #
# sum(cost$cost)
```

Each game has:

 - 4 "spiked" BoS trials where one of the rewards is high (25-30) and the other is normal (3-7)

 - ~16 normal BoS trials where both rewards are from [1-9]

 - ~10 normal ("easy") PD trials where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff). All rewards are chosen from [1-12] respecting this.

 - ~10 “sacrifice” ("hard") PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff). All rewards are chosen from [1-12] respecting this.

```{r, include=F, eval=F}
d.treatments <- read_csv(here(data_location, "treatments.csv")) %>%
  rename(treatmentId = `_id`) %>%
  select(treatmentId, name) |>
  mutate(
    chat_cond = ifelse(str_detect(name, "no chat"), "no_chat", "chat"),
    game_cond = "oct2024"
  )

@@

d.complete <- d.round_results.raw %>%
  group_by(gameId, playerId) %>%
  summarize(max_rep = max(repNum)) %>%
  filter(max_rep == 39) %>%
  select(gameId, playerId) %>%
  filter(gameId!="AQXAv4FKrZxBrTQgE") |> # the game where someone was clever and played themselves
  write_csv(here(data_location, "complete.csv"))

@@

53 games in each condition

```{r}
read_csv(here(data_location,"raw_results.csv")) |> select(chat_cond, gameId) |> unique() |> group_by(chat_cond) |> tally()

@@

exit %>%
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes))
```
mostly think partner is human, oh good.

@@

ggplot(bonuses, aes(x = chat_cond, y = payoff, color = str_c(chat_cond))) +
  geom_jitter(width = .1, alpha = .1) +
  facet_grid(. ~ gametype) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black", geom="errorbar") +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )
```

Looks like chat does better in spike, and a little bit on normal BoS?

@@

how_many_chat <- game_chat |> filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(gameId) %>%
  summarize(pct_chat = mean(is.chat)) |> 
  arrange(pct_chat)

#how_many_chat |> nrow()
#how_many_chat |> filter(pct_chat>=.1) |> nrow()

how_many_chat |> 
  mutate(x=1/n()*row_number()) |> ggplot(aes(x=x, y=pct_chat))+geom_point()+
    coord_cartesian(ylim=c(0,1))+geom_hline(yintercept=.1)+geom_hline(yintercept=.5)+labs(title="Game 5")

```
Not a lot of talking.

@@

```{r}
ggplot(game_chat, aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
  stat_summary(fun.data = "mean_cl_boot", geom="errorbar") +
    stat_summary(fun.data = "mean_cl_boot") +

  facet_grid(. ~ gametype) +
  geom_smooth(method = "lm")
```

Looks like mostly the actually using the chat is what's helping.

@@

overall_talk <- game_chat |> left_join(group_talk)

ggplot(overall_talk |> filter(chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  geom_jitter(alpha = .1) +
  geom_smooth(method = "lm") +
  facet_grid(. ~ gametype) +
  theme(legend.position = "bottom")
```

Looks like talking on *other* trials might help in BoS if you didn't talk on this specific trial? But might just be fitting to outliers? Will need models.

@@

Near chance if you can't talk, or if you don't talk, far above chance if you do coordinate.

@@

## spike BoS

```{r}
outcome %>%
  filter(gametype == "spikeBoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "Spiked BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")

outcome %>%
  filter(gametype == "spikeBoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = repNum, y = outcome_parity)) +
  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
  stat_summary(fun.data="mean_cl_boot", geom="point")+
  geom_smooth(method = "lm", se = F) +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "spike BoS:fraction choosing good outcomes", y = "Good outcome?")
```

## easy PD

@@

Can get a reasonable option no matter what.

```{r}
outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = repNum, y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot", geom="point", alpha=.5) +
  facet_grid(.~str_c(chat_cond, "_", is_chat))+
  geom_smooth(method="lm")+
  geom_hline(aes(yintercept = .5)) 
```

## hard PD

In hard PD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AB/BA is welfare maximizing.

```{r}
outcome %>%
  filter(gametype == "hardPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) 
```
So if you do use the chat, you tend to get the best option. (Using chat means that easy and hard PD look different)

Whereas if you don't chat, they look more similar at least.

```{r}
outcome %>%
  filter(gametype == "hardPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = repNum, y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot", geom="point", alpha=.5) +
  facet_grid(.~str_c(chat_cond, "_", is_chat))+
  geom_smooth(method="lm")+
  geom_hline(aes(yintercept = .5)) 
```

@@

```{r}
pair_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, playerId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    is_chat = ifelse(words > 0, 1, 0)
  ) |> 
  group_by(gameId, gametype, chat_cond, repNum) |>
  summarize(how_many_chat = sum(is_chat)) |> 
  group_by(gameId, gametype, chat_cond) |>
  summarize(pct_both = mean(how_many_chat == 2),
            pct_either=mean(how_many_chat > 0)) |> 
  filter(chat_cond == "chat") |> 
  pivot_longer(pct_both:pct_either)

ggplot(pair_chat, aes(x = gametype, y = value, color=name, group=name)) +
  geom_point(position=position_jitterdodge(dodge.width=.7,jitter.width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", color = "black", position=position_dodge(.7))
```

Even in games that talk, there aren't *that* many trials where both people talk?

<!--
# Models

```{r,eval=F}
game_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

@@

```{r, eval=F}
perf_priors <- c(
  set_prior("normal(5,2)", class = "Intercept"),
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 2)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

@@

```{r, eval=F}
for_log_words <- for_mod |>
  filter(words > 0) |>
  mutate(log_words = log(words))

@@

```{r, eval=F}
perf_word_mod <- brm(
  payoff ~ log_words * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_log_words,
  prior = perf_priors,
  file = here(model_location, "perf_words_expt1"),
  control = list(adapt_delta = .99)
)

@@

```{r, eval=F}
exit <- read_csv(here(data_location, "exit.csv")) |> select(chat_cond, strategy) |> filter(!is.na(strategy))

@@

```{r, eval=F}
color_words <- c("red","blue", "purple", "green","yellow", "brown", "pink", "violet","cyan", "grey", "peach", "purplish", "beige", "orange")
agreement_words <- c("okay", "ok", "yes", "sure")
color_regex <- str_c(color_words, collapse="|")
agreement_regex <- str_c(agreement_words, collapse="|")

@@

-->
